### 2013-02-15 doc/icml2013/algo.tex
theory done

%%% second moments
Intuitively, performing regression on only $y$ given $x$ provides only first-order
information.  The key insight is that we can perform regression
on higher order powers to obtain more information about the parameters.
Specifically, for an integer $p \ge 1$, let us define the average
$m$-th order tensor power of the parameters as follows:
\begin{align}
M_p &\eqdef \sum_{h=1}^k \pi_h \beta_h\tp{p}. \label{eqn:Mp} % \in (\Re^{d})\tp{p}.
\end{align}
Now consider performing regression on $y^2$ given $x\tp{2}$.
Expanding $y^2 = (\innerp{\beta_h}{x} + \epsilon)^2$,
using the fact that $\innerp{\beta_h}{x}^p = \innerp{\beta_h\tp{p}}{x\tp{p}}$,
we have:
\begin{align}
y^2 &= \innerp{M_2}{x\tp{2}} + \sigma^2 + \eta_2(x), \label{eqn:y2} \\
\eta_2(x) &= \innerp{\beta_h\tp{2} - M_2}{x\tp{2}} + 2 \epsilon \innerp{\beta_h\tp{2}}{x\tp{2}} + (\epsilon^2 - \sigma^2). \nonumber
\end{align}
Again, we have expressed $y^2$ has a linear function of $x\tp{2}$
with regression coefficients $M_2$, plus a known bias $\sigma^2$ and noise.\footnote{If $\sigma^2$ were not known,
we could treat it as another coefficient
to be estimated.  The coefficients $M_2$ and $\sigma^2$ can be estimated jointly
provided that $x$ does not already contain a bias ($x_j$ must be non-constant for every $j \in [d]$).}
Importantly, the noise has mean zero; 
in fact each of the three terms has mean zero
by (i) definition of $M_2$, (ii) independence of $\epsilon$ and $h$,
and (iii) the fact that $\E[\epsilon^2] = \sigma^2$.

@@

In this section, we conduct a theoretical analysis of Spectral Experts.
In particular, we prove that it enjoys consistent parameter estimates.
\begin{theorem}[Convergence of Spectral Experts]
  \label{thm:convergence}
Each $\sD_p$ (for $p = 1, 2, 3$) consists of $n$ i.i.d. points drawn from the mixture
of linear regressions model with parameter $\theta^*$
(having three independent copies simplifies the analysis).
Assume that $\E[\cvec(x\tp{p})\tp{2}] \succ 0$ for each $p \in \{1,2,3\}$.
Assume $d \ge k$.
Let $\|x\|_2 \le R$ with probability $1$
and let $\|\beta_h\|_2 \le L$ for all $h \in [k]$.
There exists universal constants $c_1,c_2$ such that
if the number of samples is
$n \ge \frac{c_1 k d^2 L R}{\epsilon}$
and $\lambda_n^{(2)} \ge \frac{c_2 L R}{\sqrt{n}}$,
and $\lambda_n^{(3)} \ge \frac{c_3 L R}{\sqrt{n}}$,
then the parameter estimates $\hat\theta$ returned by
\algorithmref{algo:spectral-experts} (with the columns appropriate permuted)
satisfies 
$\|\hat\pi - \pi^*\|_{\infty} \le \frac{\epsilon}{???}$
and for all $h \in [k]$,
$\|\hat\beta_h - \beta^*_h\|_2 \le \epsilon ???$,
\end{theorem}
\begin{proof}
Let $T^* = M_3^*(W^*, W^*, W^*)$.
By Theorem 5.1 of \cite{AnandkumarGeHsu2012},
Need to convert error from $\epsilon$
hit by $(W^\top)^\dagger$
\end{proof}

>>>>

%%% second moments
Intuitively, performing regression on only $y$ given $x$ provides only first-order
information.  The key insight is that we can perform regression
on higher order powers to obtain more information about the parameters.
Specifically, for an integer $p \ge 1$, let us define the average
$m$-th order tensor power of the parameters as follows:
\begin{align}
M_p &\eqdef \sum_{h=1}^k \pi_h \beta_h\tp{p}. \label{eqn:Mp} % \in (\Re^{d})\tp{p}.
\end{align}
Now consider performing regression on $y^2$ given $x\tp{2}$.
Expanding $y^2 = (\innerp{\beta_h}{x} + \epsilon)^2$,
using the fact that $\innerp{\beta_h}{x}^p = \innerp{\beta_h\tp{p}}{x\tp{p}}$,
we have:
\begin{align}
y^2 &= \innerp{M_2}{x\tp{2}} + \sigma^2 + \eta_2(x), \label{eqn:y2} \\
\eta_2(x) &= \innerp{\beta_h\tp{2} - M_2}{x\tp{2}} + 2 \epsilon \innerp{\beta_h}{x} + (\epsilon^2 - \sigma^2). \nonumber
\end{align}
Again, we have expressed $y^2$ has a linear function of $x\tp{2}$
with regression coefficients $M_2$, plus a known bias $\sigma^2$ and noise.\footnote{If $\sigma^2$ were not known,
we could treat it as another coefficient
to be estimated.  The coefficients $M_2$ and $\sigma^2$ can be estimated jointly
provided that $x$ does not already contain a bias ($x_j$ must be non-constant for every $j \in [d]$).}
Importantly, the noise has mean zero; 
in fact each of the three terms has mean zero
by (i) definition of $M_2$, (ii) independence of $\epsilon$ and $h$,
and (iii) the fact that $\E[\epsilon^2] = \sigma^2$.

@@

In this section, we conduct a theoretical analysis of Spectral Experts.
In particular, we prove that it enjoys consistent parameter estimates.
\begin{theorem}[Convergence of Spectral Experts]
  \label{thm:convergence}
Each $\sD_p$ (for $p = 1, 2, 3$) consists of $n$ i.i.d. points drawn from the mixture
of linear regressions model with parameter $\theta^*$
(having three independent copies simplifies the analysis).
Assume that $\E[\cvec(x\tp{p})\tp{2}] \succ 0$ for each $p \in \{1,2,3\}$
and that $d \ge k$.
Let $\|x\|_2 \le R$ with probability $1$
and let $\|\beta_h\|_2 \le L$ for all $h \in [k]$.
Suppose the number of samples is
$n = \Omega(d^p R^p \sigmamin(\Sigma_p)^{-1} \sqrt{p \log(d) \log(1/\delta)})$ and
$n = \Omega\left( \sigma^6 L^6 R^12 k \sigmamin(\Sigma_p)^{-2} \frac{\log^3 (1/\delta)}{\epsilon^2} \right)$,
and that the regularization strengths $\lambda_n^{(2)}$ and $\lambda_n^{(3)}$ are
$\Omega(\sigma^3 L^3 R^6 \sqrt{\frac{\log(1/\delta)}{n}})$,
then the parameter estimates $\hat\theta$ returned by
\algorithmref{algo:spectral-experts} (with the columns appropriate permuted)
satisfies 
$\|\hat\pi - \pi^*\|_{\infty} \le \epsilon$
and for all $h \in [k]$,
$\|\hat\beta_h - \beta^*_h\|_2 \le \frac{\epsilon}{\pi_h^*}$.
\end{theorem}

\begin{proof}
By \reflem{lowRank} along with \reflem{lowRankLower} and \reflem{lowRankUpper} (discussed below),
we can control the Frobenius norm of the parameter estimates:
If $n = \Omega(d^p R^p \sigmamin(\Sigma_p)^{-1} \sqrt{p \log(d) \log(1/\delta)})$, then
\begin{align}
  \|\hat M_p - M_p\|_F = O\left( \sigma^3 L^3 R^6 k^{\frac12} \sigmamin(\Sigma_p)^{-1} \sqrt{\frac{\log^3 (1/\delta)}{n}} \right).
  %\|\hat M_p - M_p\|_F = O\left( \frac{\sigma^3 L^3 R^6 \sqrt{k} \log^3 (1/\delta)}{\sqrt{n} (\sigmamin(\Sigma_p) - d^p R^p \sqrt{\frac{p \log(d) \log(1/\delta)}{n}}} \right).
\end{align}
The Frobenius norm upper bounds the operator norm,
so we can directly apply Theorem 5.1 of \cite{AnandkumarGeHsu2012}.
%Let $T^* = M_3^*(W^*, W^*, W^*)$.
%Need to convert error from $\epsilon$
%hit by $(W^\top)^\dagger$
\end{proof}

### 2013-02-15 doc/icml2013/regression.tex
theory done

\subsection{Low-rank Regression}
\label{sec:regression}

@@

Our analysis is based on the low-rank regression framework of
\citet{Tomioka2011} for tensors, which builds off of
\citet{NegahbanWainwright2009} for matrices.
The main challenge is to control the noise $\eta_p(x)$,
which in addition to containing the usual Gaussian part, is also dependent on
$x$, involves the mixing noise and various cross terms.
We first set up some notation to all three regressions (\refeqn{estimateM1}, \refeqn{estimateM2}, and \refeqn{estimateM3}).
Define the observation operator $\opX(M_p) : \Re^{d\tp{p}} \to \Re^{n}$
mapping compound parameters $M_p$:
\begin{align}
  \opX(M_p; \sD)_i &\eqdef \frac12 \sum_{(x,y) \in \sD} \innerp{M_p}{x\tp{p}}_i, \quad i  \in [n].
\end{align}

Let $\kappa(\opX)$ be the restricted strong constant,
and let $\opX^*(\eta_p; \sD) = \sum_{(x,y) \in \sD} \eta_p(x) x\tp{p}$
be the adjoint.

@@

In this section, we lower bound the restricted strong convexity constant
$\kappa(\opX)$.

\paragraph{Adjoint operator}

In this section, we upper bound the operator norm of the adjoint
$\|\opX(\eta_p)\|_\text{op}$.

@@

%In general, the optimization problem we'd like to solve is,
%\begin{align}
%  \hat M_p &= 
%  \arg\min_{M_p}& \frac{1}{2N} \| \vec y - \opX(M_p) \|_2^2 + \frac{\lambda_n}{K} \sum_{h=1}^K \| (M_p)_{(h)} \|_* \label{eq:regression}.
%\end{align}

\begin{lemma}[\citet{Tomioka2011}, Theorem 1]
Suppose there exists a restricted strong convexity constant $\kappa(\opX)$ such that
$$\frac{1}{2n} \| \opX( \Delta )\|_2^2 \ge \kappa(\opX) \|\Delta\|^2_F \quad \text{and} \quad
\lambda_n \ge \frac{\|\opX^*(\eta_p)\|_2}{n}.$$
Then the error of $\hat M_p$ is bounded as follows:
$\| \hat M_p - M_p^* \|_F \le \frac{\lambda_n \sqrt{k}}{\kappa(\opX)}$.
\end{lemma}

Going forward, we need to lower bound the restricted strong convexity
parameter $\kappa(\opX)$ and upper bound the adjoint operator
$\|\opX^*(\epsilon)\|_{2}^2$.
%We
%will appeal to the random design framework that models the input $x$ as
%random and show bounds that hold with high probability.

\begin{lemma}[lower bound on restricted strong convexity]
Let $\Sigma_p \eqdef \E[\cvec(x\tp{p})\tp{2}]$.
With high probability at least $1-\delta$,
for some constant $c$,
$$\kappa(\opX) \ge c \sigma_\text{min}^p(\Sigma_p)$$,
where
%$\sigma_\text{min}^p(\Sigma_p) = \min_{M_p \in \sM_p : \|M_p\|_F \le 1} \cvec(M_p)^\top \Sigma_p \cvec(M_p)$.
$\sigma_\text{min}^p(\Sigma_p) = \min_{M_p \in \sM_p : \|M_p\|_F \le 1} \trace(\cvec(M_p)^{\otimes 2} \Sigma_p)$.
\end{lemma}
Note that $\sigma_\text{min}$ 
\begin{proof}
Expanding the definition of the observation operator:
$\frac{1}{2n} \|\opX(\Delta)\|_2^2
= \frac{1}{2n} \sum_{(x,y) \in \sD} \innerp{\Delta}{x\tp{p}}^2$.
Unfolding the tensor and letting $\hat\Sigma_p \eqdef \frac{1}{n} \sum_{(x,y) \in \sD} \cvec(x\tp{p})\tp{2}$,
we have 
$\frac{1}{2n} \|\opX(\Delta)\|_2^2
\ge \frac{1}{2} \trace(\cvec(\Delta)\tp{2} \hat\Sigma_p)
\ge \frac{1}{2} \|\Delta\|_F^2 \sigma_\text{min}^p(\hat\Sigma_p)$.

%  \begin{align*}
%    \frac{\|\opX(M_p)\|_1}{N} &= \frac{1}{N}\sum_{i=1}^{N} |\innerp{x_i\tp{p}}{M_p}| \\
%    &\ge \frac{1}{N}\sum_{i=1}^{N} \sigma_{min}( M_p ) \|x_i\|^p \\
%    &\ge \frac{\kappa(M_p)}{\sqrt{D}} \frac{\sum_{i=1}^{N} \|x_i\|^p}{N} \| M_p \|_F.
%  \end{align*}
%
%  While $\kappa(M_p)$ depends on $M_p$ we can lower bound it by
%  restricting the convex space to $\{ M_p | \kappa(M_p) \ge \kappa \}$.
%  Further, using concentration bounds, we can lower bound $\|x_i\|^p$ as
%  well. \todo{What is $\E[\|x_i\|^3]$? It is some quantity greater than
%  zero}.
%  
%  Thus, 
%  \begin{align*}
%  \frac{ \|\opX(M_p)\|_2 }{2 N} 
%  &\ge
%  \frac{ \|\opX(M_p)\|_1 }{2 \sqrt{D}N} 
%  &\ge \frac{\kappa}{2 D} (\E[\|x_i\|^p] - \frac{t}{N}) \| M_p \|_F.
%  \end{align*}
%  with high probability $1 - \delta$. \todo{The bound needs to be shown
%  for $\| \opX(M_p) \|_2^2$.}
%  
%  In other words, $\kappa(\opX)
%  = \frac{\kappa}{2 D} (\E[\|x_i\|^p] - \frac{t}{N})$ is a valid
%  lower bound.
\end{proof}

\begin{lemma}[bounded error]
  Let $\opX$ be the linear operator previously defined. Then,
  $$\|\opX^*(\epsilon)\|_\text{op} \le Y$$.
\end{lemma}
\begin{proof}
%  \begin{align*}
%    \frac{\opX^*(\epsilon)}{N} 
%    &= \frac{1}{N}\sum_{i=1}^{N} \epsilon_i \innerp{x_i\tp{p}}{M_p} \\
%    &= \frac{1}{N}\sum_{i=1}^{N} (\innerp{x_i\tp{p}}{M_p - \beta_i\tp{p}} + \varepsilon)  \innerp{x_i\tp{p}}{M_p}.
%  \end{align*}
%  Let $M_p - \beta_i\tp{p} = \tilde M_i$ and consider it to be a random
%  variable.
%  \begin{align*}
%    \frac{\opX^*(\epsilon)}{N} 
%    &= \frac{1}{N}\sum_{i=1}^{N} \innerp{\tilde M_i}{x_i\tp{2p}} + \varepsilon \innerp{M_p}{x_i\tp{p}} \\
%    &\le \frac{1}{N}\sum_{i=1}^{N} \|\tilde M_i\|_{op} \|x_i\|^{2p} + \varepsilon \|M_p\|_{op}\|x_i\|^{p} \\
%    &\le \|M_p\|_{op} ( \frac{1}{N}\sum_{i=1}^{N} \|x_i\|^{2p} + \varepsilon \|x_i\|^{p} ).
%  \end{align*}
%
%  We can probabilistically bound the above by considering the union
%  bound of $\varepsilon$ and $\|x_i\|^p$.
%
%
%  Finally,
%  \begin{align*}
%    \frac{\opX^*(\epsilon)}{N} 
%    &\le \|M_p\|_{op} (\E[\|x_i\|^{2p}] + t/N + (t/N) (\E[ \|x_i\|^{p} ] + t/N) ),
%  \end{align*}
%  with high probability. Note that this does not converge to $0$ as $N
%  \to \infty$ because of the variance of the selection probability.

By Lemma 19 of \cite{hsu13spherical},
we have that $\frac{1}{n} \sum_{i=1}^n \epsilon_i^3 = O(\sqrt{\frac{\log^3(1/\delta)}{n}})$.
with probability at least $1-\delta$.

\end{proof}

>>>>

\paragraph{Low-rank Regression}
%\label{sec:regression}

@@

Our analysis is based on the low-rank regression framework of
\citet{Tomioka2011} for tensors, which builds off of
\citet{NegahbanWainwright2009} for matrices.
The main challenge is to control the noise $\eta_p(x)$,
which in addition to containing the usual Gaussian part, is also dependent on
$x$, involves the mixing noise and various cross terms.
We first set up some notation to all three regressions (\refeqn{estimateM1}, \refeqn{estimateM2}, and \refeqn{estimateM3}).
Define the observation operator $\opX_p(M_p) : \Re^{d\tp{p}} \to \Re^{n}$
mapping compound parameters $M_p$:
\begin{align}
\opX_p(M_p; \sD)_i &\eqdef \frac12 \sum_{(x,y) \in \sD} \innerp{M_p}{x\tp{p}}_i, \quad i  \in [n].
\end{align}

Let $\kappa(\opX_p)$ be the restricted strong constant,
and let $\opX^*_p(\eta_p; \sD) = \sum_{(x,y) \in \sD} \eta_p(x) x\tp{p}$
be the adjoint.

@@

In this section, we lower bound the restricted strong convexity constant
$\kappa(\opX_p)$.

@@

%In general, the optimization problem we'd like to solve is,
%\begin{align}
%  \hat M_p &= 
%  \arg\min_{M_p}& \frac{1}{2N} \| \vec y - \opX_p(M_p) \|_2^2 + \frac{\lambda_n}{K} \sum_{h=1}^K \| (M_p)_{(h)} \|_* \label{eq:regression}.
%\end{align}

\begin{lemma}[\citet{Tomioka2011}, Theorem 1]
\label{lem:lowRank}
Suppose there exists a restricted strong convexity constant $\kappa(\opX)$ such that
$$\frac{1}{2n} \| \opX_p( \Delta )\|_2^2 \ge \kappa(\opX) \|\Delta\|^2_F \quad \text{and} \quad
\lambda_n \ge \frac{\|\opX^*(\eta_p)\|_2}{n}.$$
Then the error of $\hat M_p$ is bounded as follows:
$\| \hat M_p - M_p^* \|_F \le \frac{\lambda_n \sqrt{k}}{\kappa(\opX)}$.
\end{lemma}

Going forward, we need to lower bound the restricted strong convexity
parameter $\kappa(\opX)$ and upper bound the adjoint operator
$\|\opX^*(\eta_p)\|_{2}^2$.
%We
%will appeal to the random design framework that models the input $x$ as
%random and show bounds that hold with high probability.

%\paragraph{Adjoint operator}

%In this section, we upper bound the operator norm of the adjoint
%$\|\opX_p(\eta_p)\|_\text{op}$.

First, let us lower bound the restricted strong convexity parameter $\kappa(\opX_p)$:

\begin{lemma}[lower bound on restricted strong convexity]
\label{lem:lowRankLower}
Let $\Sigma_p \eqdef \E[\cvec(x\tp{p})\tp{2}]$.
With probability at least $1-\delta$,
$$\kappa(\opX) = \Omega\left(\sigmamin(\Sigma_p) - d^p R^{p} \sqrt{\frac{p \log(d) \log (1/\delta)}{n}}\right).$$
\end{lemma}
\begin{proof}
Expanding the definition of the observation operator:
$\frac{1}{2n} \|\opX_p(\Delta)\|_2^2
= \frac{1}{2n} \sum_{(x,y) \in \sD} \innerp{\Delta}{x\tp{p}}^2$.
Unfolding the tensor and letting $\hat\Sigma_p \eqdef \frac{1}{n} \sum_{(x,y) \in \sD} \cvec(x\tp{p})\tp{2}$,
we have 
$\frac{1}{2n} \|\opX_p(\Delta)\|_2^2
\ge \frac{1}{2} \trace(\cvec(\Delta)\tp{2} \hat\Sigma_p)
\ge \frac{1}{2} \|\Delta\|_F^2 \sigmamin(\hat\Sigma_p)$.
It remains to show that $\sigmamin(\hat\Sigma_p)$ and $\sigmamin(\Sigma_p)$ are close.
To do this, we first apply Hoeffding's inequality elementwise.
Since $\|x\| \le R$, we have that for each element $(i,j)$,
$|(\hat\Sigma_p)_{ij} - (\Sigma_p)_{ij}| = O(R^{p}\sqrt{\frac{\log (1/\delta)}{n}})$.
Applying the union bound over the $d^{2p}$ elements of $\hat\Sigma_p - \Sigma_p$,
we have that the max norm is bounded:
$\|\hat\Sigma_p - \Sigma_p\|_\text{\rm max} = O(R^{p} \sqrt{\frac{p \log(d) \log (1/\delta)}{n}})$.
The max norm times $d^p$ upper bounds the Frobenius norm, which upper bounds the operator norm, so we have that
$\|\hat\Sigma_p - \Sigma_p\|_\text{\rm op} = O(d^p R^{p} \sqrt{\frac{p \log(d) \log (1/\delta)}{n}})$.
Using the fact that $\sigmamin(\hat\Sigma_p) = \sigmamin(\Sigma_p) + \|\hat\Sigma_p - \Sigma_p\|_\text{\rm op}$
yields the result.
\end{proof}

Next, let us upper bound the operator norm of the observation adjoint:

\begin{lemma}[upper bound on adjoint noise]
\label{lem:lowRankUpper}
Let $\opX_p$ be the linear operator previously defined. Then with probability at least $1-\delta$,
$$\frac1{n} \|\opX_p^*(\eta_p)\|_\op \le O\left(\sigma^3 L^3 R^6 \sqrt{\frac{\log^3(1/\delta)}{n}}\right).$$
for each $p \in \{1,2,3\}$.
\end{lemma}
\begin{proof}
Let $\hat\E_p[f(x,\epsilon,h)]$ denote the empirical expectation over the examples in dataset $\sD_p$
(recall the $\sD_p$'s are independent to simplify the analysis).
By definition,
$\frac1n \|\opX_p^*(\eta_p)\|_\op = \hat\E_p[\eta_p(x) x\tp{p}]$
for $p \in \{1,2,3\}$ defined in \refeqn{y1}, \refeqn{y2}, and \refeqn{y3}.
Each $\eta_p(x)$ is composed of several zero-mean random variables,
and since $\|A + B\|_\op \le \|A\|_\op + \|B\|_\op$,
it suffices to consider each term in turn.
For each term, we will condition on $x$, use a standard concentration inequality to
bound the noise.

First, consider $\eta_1(x)$, which contains two terms.
The first term is $\innerp{M_1 - \beta_h}{x}$.
We have that $\|M_1 - \beta_h\|_2 \le L$,
so by Hoeffding's inequality, and using the fact that $\|x\|_2 \le R$,
with probability $1-\delta_1$,
$\|\hat\E_1[\innerp{M_1 - \beta_h}{x} x]\|_\op = O(R^2 L \sqrt{\frac{\log(1/\delta_1)}{n}})$.
The second term is $\epsilon \sim \normal{0}{\sigma^2}$,
with probability $1-\delta_1$,
$\|\hat\E_1[\epsilon x]\|_\op = O(R \sigma \sqrt{\frac{\log(1/\delta_1)}{n}})$.

Now, consider $\eta_2(x)$, which contains three terms.
Using similar techniques, we have that
$\|\hat\E_2[\eta_2(x) x\tp{2}]\|_\op = O(R^2 (L^2 R^2 + \sigma L R + \sigma^2) \sqrt{\frac{\log(1/\delta_1)}{n}})$.

Finally, $\eta_3(x)$ contains five terms plus an additional bias since the estimator uses $\hat M_1$ rather than $M_1$.
There are two additional considerations.
First, to bound $\hat\E_3[\epsilon^3]$, we use Lemma 19 of \cite{hsu13spherical}.
Second, the additional bias is small by the previous bound (and it is constructed on $\sD_1$, which is independent of $\sD_3$
used to contruct $\hat M_3$.
Therefore, we incur another $O(\sigma^2 \cdot R \sigma \sqrt{\frac{\log(1/\delta_1)}{n}})$.
Handling the other terms using conventional means
$\|\hat\E_3[\eta_3(x) x\tp{3}]\|_\op = O(R^3 (L^3 R^3 + \sigma L^2 R^2 + \sigma^2 L R + \sigma^3 + \sigma^3 R) \sqrt{\frac{\log^3(1/\delta_1)}{n}})$.
Finally, taking $\delta_1 = \delta/3$, and taking the union bound,
we get our result, where we are bounding quantities fairly crudely for the sake of simplicity.
\end{proof}
